A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-02-09 paper/paper.Rmd

For `st_work` variable,  we only rename the column names from the raw, NLSY79 data. This variable is then used to calculate the next variable, `yr_workforce` for each survey round, which is the year of survey (`year`) minus the year of individual start working (`st_work`). Finally, `exp` variable is derived from the variable started with `WKSWK` in `new_data_qnames`. This variable measured since the last interview. Thus, to obtain the work experience since 1979, we calculate the cumulative value. As the measurement unit is in week, we convert this to year.

@@

Secondly, in the original dataset, wages were inflation-adjusted to 1990 prices. This is not done for the refreshed data because we plan to keep refreshing it as the data from the newest survey period release. The 1990 prices might not be relevant anymore for the future context. Hence, we keep it as the original value to allow the users to treat it as their needs.

>>>>

For `st_work` variable,  we only rename the column names from `new_data_qnames`. This variable is then used to calculate the next variable, `yr_workforce` for each survey round, which is the year of survey (`year`) minus the year of individual start working (`st_work`). Finally, `exp` variable is derived from the number of weeks worked since the last interview indicated as variable started with `WKSWK` in `new_data_qnames`. To obtain the work experience since 1979, we calculate the cumulative value. As the measurement unit is in week, we convert this to year.

@@

Secondly, in the original dataset, wages were inflation-adjusted to 1990 prices. This is not done for the refreshed data because we plan to keep refreshing it as the data from the newest survey period  release. Thus, the 1990 prices might not be relevant anymore for the future context. Hence, we keep it as the original value to allow the users to treat it as their needs.

### 2022-02-09 paper/paper.tex

For \texttt{st\_work} variable, we only rename the column names from the raw, NLSY79 data. This variable is then used to calculate the next variable, \texttt{yr\_workforce} for each survey round, which is the year of survey (\texttt{year}) minus the year of individual start working (\texttt{st\_work}). Finally, \texttt{exp} variable is derived from the variable started with \texttt{WKSWK} in \texttt{new\_data\_qnames}. This variable measured since the last interview. Thus, to obtain the work experience since 1979, we calculate the cumulative value. As the measurement unit is in week, we convert this to year.

@@

\caption{\label{tab:gender-race-table}The contingency table for gender and race for the full NLSY79 data.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Race} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-4}
Gender & Hispanic & Black & Non-Black, Non-Hispanic & Total\\
\midrule
Male & 1,000 (16\%) & 1,613 (25\%) & 3,790  (59\%) & 6,403 (100\%)\\
Female & 1,002 (16\%) & 1,561 (25\%) & 3,720  (59\%) & 6,283 (100\%)\\
\midrule
Total & 2,002 (32\%) & 3,174 (50\%) & 7,510 (118\%) & 12,686 (200\%)\\
\bottomrule
\end{tabular}
\end{table}

@@

There are two aspects of the original data that make the replication with exactly the same result is not obtained. The first one is that the calculation of the time variable provided, which is the ``work experience'', is not clearly articulated. Singer and Willett (2003) mention that the temporal variable is the years of experience since the first day of work. However, this exact same variable is not explicitly available in the database. From the NLSY79 topical guide (Bureau of Labor Statistics, U.S. Department of Labor 2021b), we find that several variables are tagged as work experience-related variables. One of them is the weeks of worked since the last interview. Hence, we use this to calculate the variable. Even though we got different result, the correlation between the original and refreshed work experience is reasonably high as shown in Figure \ref{fig:compare_xp_sw}.

Secondly, in the original dataset, wages were inflation-adjusted to 1990 prices. This is not done for the refreshed data because we plan to keep refreshing it as the data from the newest survey period release. The 1990 prices might not be relevant anymore for the future context. Hence, we keep it as the original value to allow the users to treat it as their needs.

>>>>

For \texttt{st\_work} variable, we only rename the column names from \texttt{new\_data\_qnames}. This variable is then used to calculate the next variable, \texttt{yr\_workforce} for each survey round, which is the year of survey (\texttt{year}) minus the year of individual start working (\texttt{st\_work}). Finally, \texttt{exp} variable is derived from the number of weeks worked since the last interview indicated as variable started with \texttt{WKSWK} in \texttt{new\_data\_qnames}. To obtain the work experience since 1979, we calculate the cumulative value. As the measurement unit is in week, we convert this to year.

@@

\caption{\label{tab:gender-race-table}The contingency table for gender and race for the full NLSY79 data. The percentage (rounded to closest 1\%) is out of the total corresponding to row.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Race} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-4}
Gender & Hispanic & Black & Non-Black, Non-Hispanic & Total\\
\midrule
Male & 1,000 (16\%) & 1,613 (25\%) & 3,790 (59\%) & 6,403\\
Female & 1,002 (16\%) & 1,561 (25\%) & 3,720 (59\%) & 6,283\\
\midrule
Total & 2,002 (16\%) & 3,174 (25\%) & 7,510 (59\%) & 12,686\\
\bottomrule
\end{tabular}
\end{table}

@@

There are two aspects of the original data that make the replication with exactly the same result is not obtained. The first one is that the calculation of the time variable provided, which is the ``work experience'', is not clearly articulated. Singer and Willett (2003) mention that the temporal variable is the years of experience since the first day of work. However, this exact same variable is not explicitly available in the database. From the NLSY79 topical guide (Bureau of Labor Statistics, U.S. Department of Labor 2021b), we find that several variables are tagged as work experience-related variables. One of them is the weeks of worked since the last interview. Hence, we use this to calculate the variable. Even though we got different result, the correlation between the original and refreshed work experience is reasonably high as shown in Figure \ref{fig:compare-xp-sw}.

Secondly, in the original dataset, wages were inflation-adjusted to 1990 prices. This is not done for the refreshed data because we plan to keep refreshing it as the data from the newest survey period release. Thus, the 1990 prices might not be relevant anymore for the future context. Hence, we keep it as the original value to allow the users to treat it as their needs.
